maxent model avec xvalidation (year-based test training)

### 2019-04-18 src/maxent_avec_xvalidation.Rmd
start work on binary

select mont
```{r month of interest -change}
mthna <- "apr"
mthno <- "4"
```

@@

  if (timeslice == "y2001" | timeslice == "y2006" | timeslice == "y2007"){
    betaval <- 1.75
  } else if (timeslice == "y2000"| timeslice == "y2002" | timeslice == "y2003" | timeslice == "y2004"){
    betaval <- 2
  } else if (timeslice == "y1999" | timeslice == "y2005" | timeslice == "y2008" | timeslice == "y2009"){
    betaval <- 2.25
  } else if (timeslice == "y2010" | timeslice == "y2011"){
    betaval <- 2.5
  } else if (timeslice == "y1998"){
    betaval <- 2.75
  }

@@

    for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_prev <- raster(unlayer_yr$nao_prev[td])
      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_prev) <- "nao_prev"
      nao_prev <- projectRaster(nao_prev, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      amo_sample <- raster(unlayer_yr$amo_sample[td])
      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(amo_sample) <- "amo_sample"
      amo_sample <- projectRaster(amo_sample, aea)
      amo_winter <- raster(unlayer_yr$amo_winter[td])
      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(amo_winter) <- "amo_winter"
      amo_winter <- projectRaster(amo_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      predcum <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
      dev.off() # stops automatic saving of the plot to a png
      prplotcum <- plot(predcum, png(paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".png"))) 
      dev.off() # stops automatic saving of the plot to a png
    }
}

@@

aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)

@@

asc120
asc41.18
asc1.55586
asc1045.85
asc108.03
asc11.7737
asc1151.99
asc1265.86
asc13.991
asc133.076
asc1387.38
asc147.406
asc16.5253
asc163.165
asc180.55
asc19.4298
asc199.79
asc2.66768
asc22.7576
asc221.141
asc244.891
asc26.5583
asc271.356
asc3.85628
asc30.8746
asc300.888
asc333.863
asc35.7402
asc370.689
asc411.794
asc457.626
asc47.2119
asc5.14036
asc508.64
asc53.8506
asc565.292
asc6.54303
asc61.1128
asc628.026
asc69.0217
asc697.259
asc77.6112
asc773.368
asc8.09252
asc856.679
asc86.9294
asc9.82275
asc947.448
asc97.0413

>>>>

select month
```{r month of interest -change}
mthna <- "jan"
mthno <- "1"
```

@@

  if (yr == "2003" | yr == "2004"){
    betaval <- 1
  } else if (yr == "2005"){
    betaval <- 1.5
  } else if (yr == "2006"){
    betaval <- 2.25
  }

@@

   for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      predcum <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
      dev.off() # stops automatic saving of the plot to a png

    }
}

@@

aic <- rbind(aic2003, aic2004, aic2005, aic2006)
maxr <- rbind(maxr2003, maxr2004, maxr2005, maxr2006)

@@

#binary maps

there seem to be two primary ways to create binary maps. the most common for conservation purposes seems to be to use maxsumspensspec values, however MPA also comes recommended (but is def less common). probably best to use cumulative for both (merrow 2013)

the thresholds are in the monthavgAIC.csv file in the model folder (not model/month)

```{r binary map using maxsumsenspec}
cumasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cumu_.*.asc", full.names = TRUE)
